---
title: "R Notebook"
output: html_notebook
---
```{r}
#install.packages("jsonlite")
#install.packages("maps")
#install.packages("GSODR")
#install.packages("nngeo")
#devtools::install_github('rstudio/leaflet')
#devtools::install_github('bhaskarvk/leaflet.extras')
library(nngeo)
library(maps)
library(GSODR)
library(jsonlite)
library(tidyverse)
library(gstat)
library(sp)
library(rgdal)
library(leaflet)
library(leaflet.extras)
```

```{r}
# Get a sample of cities
#cities2 <- maps::world.cities %>% 
  #filter(country.etc == 'USA') %>% 
  #arrange(desc(pop)) %>%
  #select(name, lat, long) %>% 
  #head(10)

pollen_to_clean<-read.csv("US_pollen_geographical_distribution/Data/pollen_data.csv")
pollen_clean<- pollen_to_clean %>% 
  dplyr::select(ID2,SITENAME:DepthAtLoc) %>% 
  dplyr::select(!(DepEnv:COLLECTOR))
write.csv(pollen_clean, "US_pollen_geographical_distribution/Data/pollen_clean.csv")

citiesjson<-read_json("US_pollen_geographical_distribution/Data/cities.json", simplifyVector = TRUE)
write_csv(citiesjson, "US_pollen_geographical_distribution/Data/cities.csv")

pollen_total_count <- pollen_clean %>% 
  dplyr::select(ID2:POLLSUM)
write_csv(pollen_total_count, "US_pollen_geographical_distribution/Data/pollen_total_count.csv")   

```

```{r}

cities<-read_csv("US_pollen_geographical_distribution/Data/cities.csv")
pollen_sum<-read_csv("US_pollen_geographical_distribution/Data/pollen_total_count.csv")
pollen_sum <- pollen_sum %>% 
  rename(longitude=LONDD, latitude=LATDD)

#convert to sf class 
pollen_sum_sf <-st_as_sf(pollen_sum, coords = c("longitude", "latitude"), 
                 crs = 4326)
plot(pollen_sum_sf)

cities_sf <-st_as_sf(cities, coords = c("longitude", "latitude"), 
                 crs = 4326)
plot(cities_sf)
# Get nearest stations for each city
cities_pollen = st_join(cities_sf, pollen_sum_sf, join = st_nn, k = 10, maxdist = 100000)
st_write(cities_pollen, "US_pollen_geographical_distribution/Data/cities_pollen.geojson", append = FALSE)


cities_pollen_nogeo<-cities_pollen %>% 
  data.frame() %>% 
  subset(select=c(city, state, ID2, SITENAME, ELEVATION, POLLSUM))
write.csv(cities_pollen_nogeo, "US_pollen_geographical_distribution/Data/cities_pollen_nogeo.csv")

```

# calculate avarage pollen count per city (since the the algorithm finds several stations per city and pollen count values over years per station )
```{r}
cities_pollen_nogeo<- read.csv("US_pollen_geographical_distribution/Data/cities_pollen_nogeo.csv")
Average_city_pollen<- cities_pollen_nogeo %>% 
  group_by(city, state) %>% 
  summarize(city_pollen=mean(POLLSUM, na.rm = TRUE))

Average_city_pollen %>%
  arrange(state, city, desc(city_pollen), by_group=TRUE) %>%
  dplyr::filter(state=="North Carolina") %>%
  mutate(city = fct_reorder(city, desc(city_pollen))) %>% 
  ggplot() +
           geom_col(aes(x=reorder(city, -city_pollen), y=city_pollen)) +
           theme(
             axis.text.x = element_text(
               angle = 90, vjust = 0.5, hjust=1
               )
             )
```


# converting pollen species data to sf (spatial data) and merging to cities spatial data
```{r}
#convert pollen species file to sf class 
pollen_clean <- pollen_clean %>% 
  rename(longitude=LONDD, latitude=LATDD)
pollen_spec_sf <-st_as_sf(pollen_clean, coords = c("longitude", "latitude"), crs = 4326)

# Get nearest stations for each city
cities_pollen_spec = st_join(cities_sf, pollen_spec_sf, join = st_nn, k = 100, maxdist = 100000)

# Reshaping the merged city and pollen species data into longer format

cities_pollen_spec_long <-cities_pollen_spec %>%
  pivot_longer(ABIES:XANTHIUM, names_to="species")

cities_pollen_spec_long <- cities_pollen_spec_long %>%
  mutate(species = recode(species,
                              ABIES	= "Fir", ACNEGUNA = "Box Elder", ACPENSYL = "Striped Maple", ACRUBRUM =	"Red Maple", ACSACNUM = "Silver Maple", ACSACRUM	= "Sugar Maple", ACERX	= "Maple", ALCRISPA	= "Mountain Alder", ALRUBRA	=	"Red Alder", ALRUGOSA	= "Speckled Alder", 
ALNUSX	= "Alder", AMBROSIA	= "Ragweed", AMORPHA	= "Leadplant", ANACARDI	= "Sumac Family", APIACEAE	= "Celery Family", 
AQUIFOLI	= "Holly Family", ARCEUTHOBI	=  "Mistletoe", ARECACEAE	= "Palm Family", ARMERIA	= "Thrift", ARTEMISIA	= "Sagebrush/Wormwood", 
ASTERX	= "Daisy tribe", ASTERCICHO	=  "Dandelion tribe", BETULA	= "Birch", BORAGINACEAE	= "Borage Family", BOTRYCHIUM	= "Grape Fern (spore)", BRASSICACEAE	= "Mustard Family", CACTACEAE	= "Cactus Family", CAMPANULACEA	= "Bluebells", CAPRIFOLIA	= "Honeysuckle Family", CARYA	= "Hickory", CARYOPHYLL	= "Chickweed Family", CASTANEA	= "Chestnut", CEANOTHUS	= "New Jersey Tea", CELTIS	= "Hackberry", CEPHALANTH	= "Button Bush", CERCOCAR	= "Mahogany", CHENOAMX	= "Goosefoot/Pigweed Families", CHRYSOLEP	= "Chinquapin", CORNUS	= "Dogwood", CORYLUS	= "Hazel", CUPRESSA	= "Cedar Family", CYPERACE	= "Sedge", DODECATH	= "Shooting Star", DRYAS	= "Mountain Avens", ELAEAGNX	= "Oleaster Family", EPHEDRA	= "Mexican Tea", EQUISETU	= "Horsetails", ERILEDCH	= "Leather Leaf/Labrador Tea", ERIVACCI	= "Blueberry", ERICACEX	= "Heath Family",
ERICALES	= "Ericales Order", ERIOGONUM	= "Buckwheat", EUPHORB	= "Spurge Family", FABACEAE	= "Pea Family", FAGUS	= "Beech", 
FRAXNIGR	= "Black Ash", FRAXPNAM	= "White/Green Ash", FRAXINUX	= "Ash", IVA	= "Marsh Elder", JUGCINER	= "Butternut", 
JUGNIGRA	= "Balck Walnut", JUGLANSX	=  "Walnut Family", KOENISLD	=	"Koenigia islandica/Polygonaceae", LAMIACEAE	= "Mint Family", 
LARIXPSEU	= "Larch/Douglas Fir", LARREA	= "Creosote Family", LILIACEAE	= "Lily Family", LIQUIDAM	= "Sweet Gum", LIRIODEN	= "Tulip Tree",
LYCOANNO	= "Club moss", LYCOCLAV	= "Club moss", LYCOCOMP	= "Club moss", LYCOSELA	= "Club moss", LYCOPODX	= "Club moss", 
MAGNOLIACE	= "Umbrella Tree Family", MALVACEAE	= "Mallow Family", MORACEAX	= "Mulberry Family", MYRICACX	= "Myrtle Family", 
NYSSA	= "Sour Gum", ONAGRACEAE	= "Evening Primerose Family", OSMUNDACEA	= "Royal Fern Family", OSTRYCAR	= "Hop/Hornbeam/Ironwood",
OXYRIA	= "Mountain Sorrel", PAPAVERX	= "Poppy Family", PEDICULARI	= "Louse Wort", PICGLAUC	= "White Spruce", PICMARIA	= "Black Spruce",
PICEAX	= "Spruce", PINDIPLO	= "Hard Pines", PINHAPLO	= "Soft Pines", PINUSX	= "Pine undifferentiated", PLANTAGINX	= "Plantain Family",
PLATANUS	= "Sycamore", POACEAE	= "Grass", POLEMONI	= "Phlox Family", POLYGONAX	= "Buckwheat undifferentiated", 
POLYGBIS	=	"Polygonum bistortoides", POLYGONMX	= "Knotweed", POLYGVIV	= "Bistort", POLYPOD	= "Fern Family", POPULUS	= "Poplar/Aspen", 
POTENTILLA	= "Cinquefoil", PROSOPIS	= "Mesquite", PTERIDIUM	= "Bracken", QUERCUS	= "Oak", RANUNCUL	= "Buttercup Family", 
RHAMNACEAE	= "Buckthorn/Grape", ROSACEAX	= "Rose Family", RUBIACEAE	= "Bedstraw Family", RUBUS	= "Raspberry/Cloudberry", 
RUMEX	= "Sorrel", RUMEOXYR	= "Sorrel/Mountainsorrel", SALIX	= "Willow", SANGUI	= "American Burnet", SARCOBATUS	= "Greasewood", 
SXFRACER	= "Nodding Saxifrage", SXFRAHIE	=	"Saxifraga hieracifolia", SXFRAOPP	= "Purple Saxifrage", SXFRATRI	= "Prickled Saxifrage",
SXFRAGAX	= "Saxifrage Family", SCROPHUL	= "Figwort Family", SELAGINE	= "Club moss", SHEPHERDIA	= "Soapberry", SPHAGNUM	= "Peat moss", 
TAXODIUM	= "Cypress", TAXUS	= "Yew", THALICTRUM	= "Meadow Rue", TILIA	= "Basswood", TSUGHETE	= "Western Hemlock", 
TSUGMERT	= "Mountain Hemlock", TSUGAX	= "Hemlock unifferentiated", ULMUS	= "Elm", URTICACX	= "Nettle Family", XANTHIUM	= "Cocklebur types"))

st_write(cities_pollen_spec_long, "US_pollen_geographical_distribution/Data/cities_pollen_spec_long.geojson", append = FALSE)


cities_pollen_spec_long_nogeo <- cities_pollen_spec_long %>% 
  select(city, state, ID2, SITENAME, ELEVATION, POLLSUM, YrOfSamp, DepthAtLoc,species, value)

write.csv(cities_pollen_spec_long_nogeo, "US_pollen_geographical_distribution/Data/cities_pollen_spec_long_nogeo.csv", append = FALSE)

```
#sorting pollen species by pollen count valaue for each city and select top ten species
```{r}
cities_pollen_spec_long_nogeo<-read.csv("US_pollen_geographical_distribution/Data/cities_pollen_spec_long_nogeo.csv")
cities_pollen_spec_long_nogeo %>%
  filter(value!=0| value!=NA) %>% 
  group_by(state, city) %>% 
  arrange(state, city, desc(value), by_group=TRUE) %>% 
  slice_max(order_by = value, n=10) %>% 
  filter(state=="Tennessee" & city=="Memphis") %>% 
  mutate(species = fct_reorder(species, desc(value))) %>% 
  ggplot(aes(species, value)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



#now mapping time
```{r mapWithTiles}
pollen_spec_clean <- pollen_clean %>%
  pivot_longer(ABIES:XANTHIUM, names_to="species") %>% 
  dplyr::filter(value!=0| value!=NA)

leaflet(width = 800, height = 600) %>%  
  addProviderTiles("OpenStreetMap") %>%
  addMarkers(lat = pollen_spec_clean$latitude, 
             lng = pollen_spec_clean$longitude,
             clusterOptions = markerClusterOptions(),
             popup = as.character(pollen_spec_clean$value)) %>% 
  setView(map, lng = -96,
               lat = 37.8,
               zoom = 5) 
```
```{r}
library(maps)
pollen_spec_clean$usa <- map.where("usa", pollen_spec_clean$longitude, pollen_spec_clean$latitude)

pollen_spec_clean_usa <- pollen_spec_clean %>% 
  drop_na(usa) %>% 
  arrange(species, desc(value), by_group=TRUE)


pollen_spec_clean_usa <- pollen_spec_clean_usa %>%
  mutate(species = recode(species,
                              ABIES	= "Fir", ACNEGUNA = "Box Elder", ACPENSYL = "Striped Maple", ACRUBRUM =	"Red Maple", ACSACNUM = "Silver Maple", ACSACRUM	= "Sugar Maple", ACERX	= "Maple", ALCRISPA	= "Mountain Alder", ALRUBRA	=	"Red Alder", ALRUGOSA	= "Speckled Alder", 
ALNUSX	= "Alder", AMBROSIA	= "Ragweed", AMORPHA	= "Leadplant", ANACARDI	= "Sumac Family", APIACEAE	= "Celery Family", 
AQUIFOLI	= "Holly Family", ARCEUTHOBI	=  "Mistletoe", ARECACEAE	= "Palm Family", ARMERIA	= "Thrift", ARTEMISIA	= "Sagebrush/Wormwood", 
ASTERX	= "Daisy tribe", ASTERCICHO	=  "Dandelion tribe", BETULA	= "Birch", BORAGINACEAE	= "Borage Family", BOTRYCHIUM	= "Grape Fern (spore)", BRASSICACEAE	= "Mustard Family", CACTACEAE	= "Cactus Family", CAMPANULACEA	= "Bluebells", CAPRIFOLIA	= "Honeysuckle Family", CARYA	= "Hickory", CARYOPHYLL	= "Chickweed Family", CASTANEA	= "Chestnut", CEANOTHUS	= "New Jersey Tea", CELTIS	= "Hackberry", CEPHALANTH	= "Button Bush", CERCOCAR	= "Mahogany", CHENOAMX	= "Goosefoot/Pigweed Families", CHRYSOLEP	= "Chinquapin", CORNUS	= "Dogwood", CORYLUS	= "Hazel", CUPRESSA	= "Cedar Family", CYPERACE	= "Sedge", DODECATH	= "Shooting Star", DRYAS	= "Mountain Avens", ELAEAGNX	= "Oleaster Family", EPHEDRA	= "Mexican Tea", EQUISETU	= "Horsetails", ERILEDCH	= "Leather Leaf/Labrador Tea", ERIVACCI	= "Blueberry", ERICACEX	= "Heath Family",
ERICALES	= "Ericales Order", ERIOGONUM	= "Buckwheat", EUPHORB	= "Spurge Family", FABACEAE	= "Pea Family", FAGUS	= "Beech", 
FRAXNIGR	= "Black Ash", FRAXPNAM	= "White/Green Ash", FRAXINUX	= "Ash", IVA	= "Marsh Elder", JUGCINER	= "Butternut", 
JUGNIGRA	= "Balck Walnut", JUGLANSX	=  "Walnut Family", KOENISLD	=	"Koenigia islandica/Polygonaceae", LAMIACEAE	= "Mint Family", 
LARIXPSEU	= "Larch/Douglas Fir", LARREA	= "Creosote Family", LILIACEAE	= "Lily Family", LIQUIDAM	= "Sweet Gum", LIRIODEN	= "Tulip Tree",
LYCOANNO	= "Club moss", LYCOCLAV	= "Club moss", LYCOCOMP	= "Club moss", LYCOSELA	= "Club moss", LYCOPODX	= "Club moss", 
MAGNOLIACE	= "Umbrella Tree Family", MALVACEAE	= "Mallow Family", MORACEAX	= "Mulberry Family", MYRICACX	= "Myrtle Family", 
NYSSA	= "Sour Gum", ONAGRACEAE	= "Evening Primerose Family", OSMUNDACEA	= "Royal Fern Family", OSTRYCAR	= "Hop/Hornbeam/Ironwood",
OXYRIA	= "Mountain Sorrel", PAPAVERX	= "Poppy Family", PEDICULARI	= "Louse Wort", PICGLAUC	= "White Spruce", PICMARIA	= "Black Spruce",
PICEAX	= "Spruce", PINDIPLO	= "Hard Pines", PINHAPLO	= "Soft Pines", PINUSX	= "Pine undifferentiated", PLANTAGINX	= "Plantain Family",
PLATANUS	= "Sycamore", POACEAE	= "Grass", POLEMONI	= "Phlox Family", POLYGONAX	= "Buckwheat undifferentiated", 
POLYGBIS	=	"Polygonum bistortoides", POLYGONMX	= "Knotweed", POLYGVIV	= "Bistort", POLYPOD	= "Fern Family", POPULUS	= "Poplar/Aspen", 
POTENTILLA	= "Cinquefoil", PROSOPIS	= "Mesquite", PTERIDIUM	= "Bracken", QUERCUS	= "Oak", RANUNCUL	= "Buttercup Family", 
RHAMNACEAE	= "Buckthorn/Grape", ROSACEAX	= "Rose Family", RUBIACEAE	= "Bedstraw Family", RUBUS	= "Raspberry/Cloudberry", 
RUMEX	= "Sorrel", RUMEOXYR	= "Sorrel/Mountainsorrel", SALIX	= "Willow", SANGUI	= "American Burnet", SARCOBATUS	= "Greasewood", 
SXFRACER	= "Nodding Saxifrage", SXFRAHIE	=	"Saxifraga hieracifolia", SXFRAOPP	= "Purple Saxifrage", SXFRATRI	= "Prickled Saxifrage",
SXFRAGAX	= "Saxifrage Family", SCROPHUL	= "Figwort Family", SELAGINE	= "Club moss", SHEPHERDIA	= "Soapberry", SPHAGNUM	= "Peat moss", 
TAXODIUM	= "Cypress", TAXUS	= "Yew", THALICTRUM	= "Meadow Rue", TILIA	= "Basswood", TSUGHETE	= "Western Hemlock", 
TSUGMERT	= "Mountain Hemlock", TSUGAX	= "Hemlock unifferentiated", ULMUS	= "Elm", URTICACX	= "Nettle Family", XANTHIUM	= "Cocklebur types"))

write.csv(pollen_spec_clean_usa, "US_pollen_geographical_distribution/Data/pollen_spec_clean_usa.csv")

```

```{r}
read.csv("US_pollen_geographical_distribution/Data/pollen_spec_clean_usa.csv")
leaflet(width = 800, height = 600) %>%  
  addProviderTiles("OpenStreetMap") %>%
  addAwesomeMarkers(lat = pollen_spec_clean_usa$latitude, 
             lng = pollen_spec_clean_usa$longitude,
             clusterOptions = markerClusterOptions(),
             icon = awesomeIcons(icon = "flower-sharp", markerColor = 'green'),
             popup = as.character(pollen_spec_clean_usa$species)) %>%
  setView(map, lng = -96,
               lat = 37.8,
               zoom = 4) 
```
#map species 
```{r}

leaflet(width = 800, height = 600) %>%  
  addProviderTiles("OpenStreetMap") %>%
  addAwesomeMarkers(lat = pollen_spec_clean_usa$latitude, 
             lng = pollen_spec_clean_usa$longitude,
             clusterOptions = markerClusterOptions(),
             icon = awesomeIcons(icon = "flower-sharp", markerColor = 'green'),
             popup = paste0(pollen_spec_clean_usa$species,"<br/>Discharge: ", pollen_spec_clean_usa$value)) %>%
  setView(map, lng = -96,
               lat = 37.8,
               zoom = 4) 
```
#mapping with heatmap
```{r}
leaflet(width = 800, height = 600) %>%  
  addProviderTiles("OpenStreetMap") %>%
  addHeatmap(lat = pollen_spec_clean_usa$latitude, 
             lng = pollen_spec_clean_usa$longitude,
             intensity=pollen_spec_clean_usa$value,
             blur = 20, max = 0.05, radius = 20) %>%
  setView(map, lng = -96,
               lat = 37.8,
               zoom = 4) 

```
